---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressWarnings(suppressMessages(source(here::here("/R/manifest.R"))))
source(here("R/map-theme.R"))

```
# Defining Geographic Markets for Health Care Services 

The objective of this repository and README is to lay out some thoughts, analytics, and data for defining geographic markets for health care services in the U.S.

# Mapping Geographic Markets 

The maps below show various geographic market definitions that appear in Tennessee. 

```{r,echo = FALSE}
df_hrr_map <- read_rds(here("output/tidy-mapping-files/hrr/","df_hrr.rds"))
p_hrr <- 
  df_hrr_map %>%
    filter(hrrstate %in% c("TN")) %>% 
    tbl_df() %>%
    mutate(test = factor(sample(1:10,nrow(.),replace=TRUE))) %>%
    ggplot() +
    aes(long,lat,group=group) +
    geom_polygon(aes(fill = test)) +
    geom_path(color="black") +
    coord_equal() +
  ggthemes::theme_tufte(base_family = "Gill Sans") +
  theme(legend.position = "none") +
  remove_all_axes +
  ggtitle("Hospital Referral Region\n(HRR)" ) 


df_hsa_map <- 
  read_rds(here("output/tidy-mapping-files/hsa/","df_hsa.rds")) 
p_hsa <- 
  df_hsa_map %>%
    filter(hsastate %in% c("TN")) %>% 
    tbl_df() %>%
    mutate(test = factor(sample(1:10,nrow(.),replace=TRUE))) %>%
    ggplot() +
    aes(long,lat,group=group) +
    geom_polygon(aes(fill = test)) +
    geom_path(color="black") +
    coord_equal() +
  ggthemes::theme_tufte(base_family = "Gill Sans") +
  theme(legend.position = "none") +
  remove_all_axes +
  ggtitle("Hospital Service Area\n(HSA)") 

df_pcsa_map <- read_rds(here("output/tidy-mapping-files/pcsa/","df_pcsa.rds"))
p_pcsa <- 
  df_pcsa_map %>%
    filter(pcsa_st %in% c("TN")) %>% 
    filter(id != "" & !is.na(id)) %>% 
    tbl_df() %>%
    mutate(test = factor(sample(1:10,nrow(.),replace=TRUE))) %>%
    ggplot() +
    aes(long,lat,group=group) +
    geom_polygon(aes(fill = test)) +
    geom_path(color="black") +
    coord_equal() +
  ggthemes::theme_tufte(base_family = "Gill Sans") +
  theme(legend.position = "none") +
  remove_all_axes +
  ggtitle("Primary Care Service Area\n(PCSA)") 

df_cz_map <- read_rds(here("output/tidy-mapping-files/commuting-zone/","df_cz.rds"))


county_to_cz <- readxl::read_xls(here("public-data/shape-files/commuting-zones/cz00_eqv_v1.xls")) %>% 
  janitor::clean_names()  %>% 
  rename(fips_code = fips) 

tn_czs <- 
  county_to_cz %>% mutate(commuting_zone_id_2000 = paste0(commuting_zone_id_2000)) %>% filter(str_sub(fips_code,1,2) %in% c("47")) %>% pull(commuting_zone_id_2000)


p_cz <- 
  df_cz_map %>%
  filter(commuting_zone_id_2000 %in% tn_czs) %>% 
  tbl_df() %>%
  mutate(test = factor(sample(1:10,nrow(.),replace=TRUE))) %>%
  ggplot() +
  aes(long,lat,group=group) +
  geom_polygon(aes(fill = test)) +
  geom_path(color="black") +
  coord_equal() +
  ggthemes::theme_tufte(base_family = "Gill Sans") +
  theme(legend.position = "none") +
  remove_all_axes +
  ggtitle("Commuting Zone\n(CZ-2000)") 
  

p_hrr + p_hsa + p_pcsa + p_cz + plot_layout(nrow=2)
```

# File Descriptions

## Geographic Mapping Files

- The file [R/construct-county-map-data.R](R/construct-county-map-data.md) constructs ggplot-friendly mapping data for U.S. counties (`output/tidy-mapping-files/county/df_county.rds`). It also extracts contiguous counties and estimates county centroids (`output/tidy-mapping-files/county/df_county_info.rds`).

- The file [R/construct-dartmouth-geography-data.R](R/construct-dartmouth-geography-data.R) constructs ggplot-friendly mapping data for Dartmouth Atlas geographies including Hospital Referral Region (HRR), Hospital Service Region (HSA) and Primary Care Service Region (PCSA). It also constructs a data file listed at the geographic market level, and which contains data on centroids and contiguous markets. Note that the basis for these files is the ZIP to HRR/HSA and PCSA crosswalk files available for download at the [NBER website](https://www.nber.org/data/dartmouth-atlas-geography.html) and in [archived versions of the Dartmouth Atlas webpage](http://archive.dartmouthatlas.org/tools/downloads.aspx?tab=42).  There are also shapefiles already constructed and [available for use on the archived Dartmouth website](http://archive.dartmouthatlas.org/tools/downloads.aspx?tab=39)--though I found these difficult to work with (e.g., I could not easily extract contiguous geographies using them, as I can by building up a HRR/HSA/PCSA shapefile from a ZCTA map)

- The file [R/construct-commuting-zone-data.R](R/construct-commuting-zone-data.R) constructs ggplot-friendly mapping data for Commuting Zones defined using the 2000 census. The underying county-to-commuting zone data can be found on the [USDA website](https://www.ers.usda.gov/data-products/commuting-zones-and-labor-market-areas/). 

## Patient Sharing Files

- The file [R/read-and-tidy-cms-hospital-service-areas.R](R/read-and-tidy-cms-hospital-service-areas.R) reads in the CMS Hospital Service Area file for 2017. Note the source of these data are downloaded csv files from the interactive CMS data explorer available at the links at [this link](https://www.cms.gov/Research-Statistics-Data-and-Systems/Statistics-Trends-and-Reports/Hospital-Service-Area-File/index.html). The final file is rolled up to the FIPS county level and is storeed in `output/hospital-county-patient-data/2017/hospital-county-patient-data.rds`. 


